---
title: "Peskett_et_al_2026_Data-Analysis-and-Bioinformatics"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Set-Up

Load required packages:

```{r}
library(tidyverse)
library(car)
library(lme4)
library(lmerTest)
library(DHARMa)
library(lattice)
library(ggthemes) # provides a colour-blind safe palette
library(sjPlot)
library(decontam)
library(qiime2R)
library(phyloseq)
library(gridExtra)
library(vegan)
library(ggpubr)
library(gplots)
library(corrplot)
library(Hmisc)
library(microViz) # to add variables into phyloseq object
library(plyr)
library(data.table)
library(glue)
library(ggh4x) #packages for making nested facets
library(ggtext) #needed for markdown elements
library(readxl)
library(glue)
library(reshape2)
library(lubridate)
library(plotly)
library(ggpmisc)
library(SQMtools)
library(Biostrings)
library(rlang)
library(patchwork)
library(wesanderson)
library(pheatmap)
```

# Chemistry Analyses and Visualizations

```{r}
data <- read.csv("Peskett_ChemistryData.csv")
```

Filter data:

Only include biotic incubations and PFAS added bottles

```{r}
chemdata <- data %>% 
  dplyr::filter(PFAS !="No") %>% 
  dplyr::filter(Incubation != "Abiotic") 

#Make pH a categorical variable:
chemdata$pH <- as.factor(chemdata$pH)
```

### diPAP

```{r}
diPAP <- chemdata %>% 
  select(1:9, 11) %>% 
  na.omit()

diPAP1 <- lm(diPAP ~ Participant + pH + Redox + Carbon + Time, data = diPAP)
summary(diPAP1)

#Check model assumptions:

par(mfrow = c(2,2))
plot(diPAP1) # Variance increasing 

#Try a log-transformation maybe

diPAP$logdiPAP <- log10(diPAP$diPAP)

diPAP2 <- lm(logdiPAP ~ Participant + pH + Redox + Carbon + Time, data = diPAP)
summary(diPAP2)

plot(diPAP2) 

```

Interaction - pH and diPAP

```{r}
diPAP3 <- lm(diPAP ~ pH*Time, data = diPAP)
summary(diPAP3)

par(mfrow = c(2,2))
plot(diPAP3)

#Variance increasing - log transform

diPAP4 <- lm(logdiPAP ~ pH*Time, data = diPAP)
summary(diPAP4)

plot(diPAP4)

```

## monoPAP

```{r}
monoPAP <- chemdata %>% 
  select(1:9, 12) %>% 
  na.omit

monoPAP1 <- lm(monoPAP ~ Participant + pH + Redox + Carbon + Time, data = monoPAP)
summary(monoPAP1)

#Check model assumptions:

par(mfrow = c(2,2))
plot(monoPAP1)
```

## FTOH

Non-parametric

Wilcox t-tests for pH, participant, Redox, and Carbon:

```{r}
FTOH <- chemdata %>% 
  select(1:9, 13) %>% 
  na.omit

# Participant

participant <- wilcox.test(FTOH ~ Participant, data = FTOH, exact = FALSE)
participant

# pH

pH <- wilcox.test(FTOH ~ pH, data = FTOH, exact = FALSE)
pH

# Redox

redox <- wilcox.test(FTOH ~ Redox, data = FTOH, exact = FALSE)
redox

# Carbon

carbon <- wilcox.test(FTOH ~ Carbon, data = FTOH, exact = FALSE)
carbon
```

Spearman-rank correlation for time:

```{r}
time <- cor.test(FTOH$FTOH, FTOH$Time,  method = "spearman")
time
```

## PFHxA

Non-parametric

Wilcox t-tests for pH, participant, Redox, and Carbon:

```{r}
PFHxA <- chemdata %>% 
  select(1:9, 14) %>% 
  na.omit

# Participant

participant2 <- wilcox.test(PFHxA ~ Participant, data = PFHxA, exact = FALSE)
participant2

# pH

pH2 <- wilcox.test(PFHxA ~ pH, data = PFHxA, exact = FALSE)
pH2

# Redox

redox2 <- wilcox.test(PFHxA ~ Redox, data = PFHxA, exact = FALSE)
redox2

# Carbon

carbon2 <- wilcox.test(PFHxA ~ Carbon, data = PFHxA, exact = FALSE)
carbon2
```

Spearman-rank correlation for time:

```{r}
time2 <- cor.test(PFHxA$PFHxA, PFHxA$Time,  method = "spearman")
time2
```

## Visualizations

### Parent Compounds

#### Abiotic

```{r}
NormalizedAbiotic <- data %>% 
  select(1:9, 15:16) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  pivot_longer(cols = 10:11,
               names_to = "Analyte",
               values_to = "Concentration") %>% 
  filter(Time %in% c("0", "24", "168")) %>% 
  filter(Incubation == "Abiotic") %>% 
  na.omit() %>% 
  mutate(Analyte = case_when(
    Analyte == "PFOA_Norm" ~ "PFOA",
    Analyte == "diPAP_Norm" ~ "diPAP"
  ))
```

```{r}
NormalizedAbiotic$Analyte <- factor(NormalizedAbiotic$Analyte,
                       levels = c("PFOA", "diPAP"))

plot1 <- ggplot(NormalizedAbiotic, aes(x = Time, y = Concentration, color = Analyte, group = interaction(Analyte,Participant,Bottle_ID))) +
  geom_line(size = 1, colour = "grey") +
  geom_point(size = 3.5, colour = "blue") +
  labs(x = "Time (h)", y = "Normalized Concentration (% of T0)") +
  facet_grid(~Analyte,
             scales = "free_y",
              labeller = labeller(
      Analyte = c(
        PFOA = "PFOA",
        diPAP = "6:2 diPAP"
      ))) +
  scale_x_continuous(
  breaks = c(0, 24, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 20, face = "bold")) +
  theme(axis.title.y = element_text(size = 20, face = "bold")) +
  theme(strip.text.x = element_text(size = 17, face = "bold")) +
  theme(strip.text.y = element_text(size = 17, face = "bold")) +
  theme(axis.text.y = element_text(size = 17)) + 
  theme(axis.text.x = element_text(size = 17)) +
  theme(legend.position="none")

plot1
```

#### Biotic

```{r}
NormalizedBiotic <- data %>% 
  select(1:9, 15:16) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  pivot_longer(cols = 10:11,
               names_to = "Analyte",
               values_to = "Concentration") %>% 
  filter(Time %in% c("0","72", "168")) %>% 
  filter(Incubation == "Biotic") %>% 
  na.omit() %>% 
  mutate(Analyte = case_when(
    Analyte == "PFOA_Norm" ~ "PFOA",
    Analyte == "diPAP_Norm" ~ "diPAP"
  ))
```

```{r}
NormalizedBiotic$Analyte <- factor(NormalizedBiotic$Analyte,
                       levels = c("PFOA", "diPAP"))

plot2 <- ggplot(NormalizedBiotic, aes(x = Time, y = Concentration, color = Analyte, group = interaction(Analyte,Participant,Bottle_ID))) +
  geom_line(size = 1, colour = "grey") +
  geom_point(size = 4, colour = "blue") +
  labs(x = "Time (h)", y = "Normalized Concentration (% of T0)") +
  facet_grid(~Analyte,
             scales = "free_y",
              labeller = labeller(
      Analyte = c(
        PFOA = "PFOA",
        diPAP = "6:2 diPAP"
      ))) +
  scale_x_continuous(
  breaks = c(0, 72, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 20, face = "bold")) +
  theme(axis.title.y = element_text(size = 20, face = "bold")) +
  theme(strip.text.x = element_text(size = 17, face = "bold")) +
  theme(strip.text.y = element_text(size = 17, face = "bold")) +
  theme(axis.text.y = element_text(size = 17)) + 
  theme(axis.text.x = element_text(size = 17)) +
  theme(legend.position="none")

plot2

```

### Products

#### MonoPAP

```{r}
monoPAP <- data %>% 
  select(1:9, 12) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Biotic")
```

Summary statistics:

```{r}
monoPAP_summary <- monoPAP %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(monoPAP, na.rm = TRUE),
    se = sd(monoPAP, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

Biotic:

```{r}
monoPAP_summary$pH <- as.factor(monoPAP_summary$pH)

plot3 <- ggplot(monoPAP_summary, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 4) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
    scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL)", title = "6:2 monoPAP") +
   scale_x_continuous(
  breaks = c(0, 4, 8, 24, 72, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  geom_hline(
    yintercept = 1.58,
    linetype = "dashed"
  ) +
  facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
  theme(plot.title = element_text(size = 15, face = "bold")) +
  theme(legend.text = element_text(size = 17),   # Increases legend item text
    legend.title = element_text(size = 20)  # Increases legend title text
  )

plot3

```

Abiotic:

```{r}
monoPAP2 <- data %>% 
  select(1:9, 12) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Abiotic")
```

Summary statistics:

```{r}
monoPAP_summary2 <- monoPAP2 %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(monoPAP, na.rm = TRUE),
    se = sd(monoPAP, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

By Participant:

```{r}
monoPAP_summary2$pH <- as.factor(monoPAP_summary2$pH)

plot3.2 <- ggplot(monoPAP_summary2, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 3.5) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
    scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL)", title = "6:2 monoPAP") +
   scale_x_continuous(
  breaks = c(0, 24, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  geom_hline(
    yintercept = 1.58,
    linetype = "dashed"
  ) +
  facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
   theme(plot.title = element_text(size = 15, face = "bold"))

plot3.2

```

#### FTOH

Biotic:

```{r}
FTOH <- data %>% 
  select(1:9, 13) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Biotic")
```

Summary statistics:

```{r}
FTOH_summary <- FTOH %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(FTOH, na.rm = TRUE),
    se = sd(FTOH, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

```{r}
FTOH_summary$pH <- as.factor(FTOH_summary$pH)

plot4 <- ggplot(FTOH_summary, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 4) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
  scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL)", title = "6:2 FTOH") +
   scale_x_continuous(
  breaks = c(0, 4, 8, 24, 72, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  geom_hline(
    yintercept = 1.60,
    linetype = "dashed"
  ) +
   facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
  theme(plot.title = element_text(size = 15, face = "bold")) +
    theme(strip.background = element_blank())

plot4

```

Abiotic: 

Abiotic

pH x Participant

```{r}
FTOH2 <- data %>% 
  select(1:9, 13) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Abiotic")
```


```{r}
FTOH_summary2 <- FTOH2 %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(FTOH, na.rm = TRUE),
    se = sd(FTOH, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

```{r}
FTOH_summary2$pH <- as.factor(FTOH_summary2$pH)

plot4.2 <- ggplot(FTOH_summary2, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 3.5) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
  scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL fecal susp.)", title = "6:2 FTOH") +
   scale_x_continuous(
  breaks = c(0, 4, 8, 24, 72, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off", ylim = c(0, 0.05)) +
   geom_hline(
    yintercept = 1.60,
    linetype = "dashed"
  ) +
   facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
  theme(plot.title = element_text(size = 15, face = "bold")) +
    theme(strip.background = element_blank())

plot4.2
```

#### PFHxA

Biotic:

```{r}
PFHxA <- data %>% 
  select(1:9, 14) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Biotic")
```

Summary statistics:

```{r}
PFHxA_summary <- PFHxA %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(PFHxA, na.rm = TRUE),
    se = sd(PFHxA, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

```{r}
PFHxA_summary$pH <- as.factor(PFHxA_summary$pH)

plot5 <- ggplot(PFHxA_summary, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 4) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
    scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL)", title = "PFHxA") +
   scale_x_continuous(
  breaks = c(0, 4, 8, 24, 72, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  geom_hline(
    yintercept = 1.33,
    linetype = "dashed"
  ) +
  facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
   theme(plot.title = element_text(size = 15, face = "bold")) +
    theme(strip.background = element_blank())

plot5

```

Abiotic:

```{r}
PFHxA2 <- data %>% 
  select(1:9, 14) %>% 
  filter(PFAS !="No") %>% #Filer out no PFAS bottles
  na.omit() %>% 
  filter(Incubation == "Abiotic")
```

Summary statistics:

```{r}
PFHxA_summary2 <- PFHxA2 %>%
  dplyr::group_by(Time, pH, Participant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean = mean(PFHxA, na.rm = TRUE),
    se = sd(PFHxA, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )
```

```{r}
PFHxA_summary2$pH <- as.factor(PFHxA_summary2$pH)

plot5.2 <- ggplot(PFHxA_summary2, aes(x = Time, y = mean, shape = pH, colour = pH)) +
  geom_point(size = 3.5) +
    geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                width = 2,
                linewidth = 0.6) +
    scale_color_manual(values = c("2" = "black",
                                "7" = "hotpink")) +
  labs(x = "Time (h)", y = "Concentration (ng/mL)", title = "PFHxA") +
   scale_x_continuous(
  breaks = c(0, 24, 168),
  expand=expand_scale(mult=c(0,0.1))) +
  coord_cartesian(clip = "off") +
  geom_hline(
    yintercept = 1.33,
    linetype = "dashed"
  ) +
  facet_grid(~Participant,
             labeller = labeller(
    Participant = c ("A" = "Participant A",
                     "B" = "Participant B"))) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 15, face = "bold")) +
  theme(axis.title.y = element_text(size = 15, face = "bold")) +
  theme(axis.text.y = element_text(size = 14)) + 
  theme(axis.text.x = element_text(size = 14)) +
  theme(strip.text.x = element_text(size = 13, face = "bold")) +
   theme(plot.title = element_text(size = 15, face = "bold")) +
    theme(strip.background = element_blank())

plot5.2
```

#### Combine

Biotic:

```{r}
plot3_clean <- plot3 +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank()
  )

plot4_clean <- plot4 +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )

plot5_clean <- plot5 +
  theme(
   axis.title.x = element_blank(),
   axis.title.y = element_blank(),
   strip.background = element_blank(),
   strip.text.x = element_blank()
  )

Products <- ggarrange(
  plot3_clean, plot4_clean, plot5_clean,
  labels = c("A)", "B)", "C)"),
  ncol = 1, nrow = 3,
  common.legend = TRUE,
  legend = "top",
  align = "v"
)

Products <- annotate_figure(
  Products,
  left = text_grob(
    "Concentration (ng/mL)",
    rot = 90,
    size = 20,
    face = "bold"
  ),
  bottom = text_grob(
    "Time (h)",
    size = 20,
    face = "bold"
  )
)

Products

```

Abiotic:

```{r}
Products2 <- ggarrange(plot3.2, plot4.2, plot5.2,
          labels = c("A)", "B)", "C)"),
          ncol = 1, nrow = 3,
          common.legend = TRUE, 
          legend = "bottom")
Products2

plot3.2_clean <- plot3.2 +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_blank()
  )

plot4.2_clean <- plot4.2 +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )

plot5.2_clean <- plot5.2 +
  theme(
   axis.title.x = element_blank(),
   axis.title.y = element_blank(),
   strip.background = element_blank(),
   strip.text.x = element_blank()
  )

Products2 <- ggarrange(
  plot3.2_clean, plot4.2_clean, plot5.2_clean,
  labels = c("A)", "B)", "C)"),
  ncol = 1, nrow = 3,
  common.legend = TRUE,
  legend = "top",
  align = "v"
)

Products2 <- annotate_figure(
  Products2,
  left = text_grob(
    "Concentration (ng/mL)",
    rot = 90,
    size = 15,
    face = "bold"
  ),
  bottom = text_grob(
    "Time (h)",
    size = 15,
    face = "bold"
  )
)

Products2
```

# Ampliseq - greengenes2 classifer

## Run ampliseq pipeline 

```{bash}
nextflow run nf-core/ampliseq \
   -r 2.11.0 \
   -profile singularity \
   --input "Peskett_Microbiome_Samplesheet.csv" \
   --FW_primer GTGYCAGCMGCCGCGGTAA \
   --RV_primer GGACTACNVGGGTWTCTAAT \
   --metadata "Peskett_Microbiome_Metadata.txt" \
   --trunclenf 270 \
   --trunclenr 270 \
   --skip_dada_taxonomy \
   --qiime_ref_taxonomy greengenes2 \
   --save_intermediates TRUE \
   --ancom TRUE \
   --ancombc TRUE \
   --outdir "./Peskett_Ampliseq_Results"
```

# Decontam with qiime2

## Set-up

```{bash}
conda activate qiime2-amplicon-2024.10
```

## decontam-identify

Running decontam using the prevalence method:

```{bash}
qiime quality-control decontam-identify \
--i-table filtered-table.qza \
--m-metadata-file Peskett_Microbiome_Metadata_decontam.txt \
--p-method prevalence \
--p-prev-control-column Sample_or_Control \
--p-prev-control-indicator control \
--o-decontam-scores prev_decontam_scores.qza
```

## decontam-score-viz

```{bash}
qiime quality-control decontam-score-viz \
--i-decontam-scores prev_decontam_scores.qza \
--i-table filtered-table.qza \
--i-rep-seqs filtered-sequences.qza \
--p-threshold 0.5 \
--p-no-weighted \
--p-bin-size 0.05 \
--o-visualization decontam_score_viz.qzv
```

## decontam-remove

```{bash}
qiime quality-control decontam-remove \
--i-decontam-scores prev_decontam_scores.qza \
--i-table filtered-table.qza \
--i-rep-seqs filtered-sequences.qza \
--p-threshold 0.5 \
--o-filtered-table clean-table.qza \
--o-filtered-rep-seqs clean-rep-seqs.qza
```

# Remove control samples from cleaned data

Some features belonging to control samples remain because they do not show up in the samples. Therefore, going to filter the data to remove remaining control samples because we are not interested at this moment in analyzing those further. 

Requires metadata file where the first column in the file contains the IDs that should be retained titled "samples-to-keep.txt".

Filter out extraction and sequencing blanks:

```{bash}
qiime feature-table filter-samples \
  --i-table clean-table.qza \
  --m-metadata-file samples-to-keep.txt \
  --o-filtered-table id-filtered-table.qza
```

Filter sequences based on new filtered feature table:

```{bash}
qiime feature-table filter-seqs \
  --i-data clean-rep-seqs.qza \
  --i-table id-filtered-table.qza \
  --o-filtered-data clean-filtered-rep-seqs.qza
```

# Generate Feature Table Summaries

```{bash}
qiime feature-table summarize \
  --i-table id-filtered-table.qza \
  --o-visualization clean-filtered-table.qzv \
  --m-sample-metadata-file Peskett_Microbiome_Metadata_decontam.txt
  
qiime feature-table tabulate-seqs \
  --i-data clean-filtered-rep-seqs.qza \
  --o-visualization clean-filtered-rep-seqs.qzv
```

# Generate a tree for phylogenetic diversity analyses

```{bash}
qiime phylogeny align-to-tree-mafft-fasttree \
  --i-sequences clean-filtered-rep-seqs.qza \
  --o-alignment aligned-clean-filtered-rep-seqs.qza \
  --o-masked-alignment masked-aligned-clean-filtered-rep-seqs.qza \
  --o-tree unrooted-tree.qza \
  --o-rooted-tree rooted-tree.qza
```

# Alpha and beta diversity analysis

## Compute core metrics

Need to set sampling-depth using info from feature table - going to set sampling depth to 17,577 since that is the lowest frequency in our samples - sample B8F12


```{bash}
qiime diversity core-metrics-phylogenetic \
  --i-phylogeny rooted-tree.qza \
  --i-table id-filtered-table.qza \
  --p-sampling-depth 17577 \
  --m-metadata-file Peskett_Microbiome_Metadata_decontam.txt \
  --output-dir core-metrics-results
```

Next we will explore the microbial composition of the samples in the context of the sample metadata.

## Test for associations between categorical metadata columns and alpha diversity data. 

Import metadata file for analyses (i.e. does not contain blanks) "Peskett_Microbiome_Metadata_NoBlanks.txt"

```{bash}
qiime diversity alpha-group-significance \
  --i-alpha-diversity core-metrics-results/shannon_vector.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --o-visualization shannon-group-significance.qzv
```

## Test for correlations between alpha diversity and continuous variables

```{bash}

qiime diversity alpha-correlation \
  --i-alpha-diversity core-metrics-results/shannon_vector.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --o-visualization shannon-correlation-significance.qzv
```

# Analyze sample composition in the context of categorical metadata using PERMANOVA 

## Unweighted UniFrac distances.

Can pick specific columns in the metadata to analyze:

```{bash}

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Participant \
  --p-pairwise \
  --o-visualization unweighted-unifrac-participant-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column pH \
  --p-pairwise \
  --o-visualization unweighted-unifrac-pH-group-significance.qzv
  
qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column PFAS \
  --p-pairwise \
  --o-visualization unweighted-unifrac-PFAS-group-significance.qzv
  
qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Oxygen \
  --p-pairwise \
  --o-visualization unweighted-unifrac-oxygen-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Sugar \
  --p-pairwise \
  --o-visualization unweighted-unifrac-sugar-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Bottle \
  --p-pairwise \
  --o-visualization unweighted-unifrac-bottle-group-significance.qzv
```

## Weighted UniFrac

```{bash}

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Participant \
  --p-pairwise \
  --o-visualization weighted-unifrac-participant-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column pH \
  --p-pairwise \
  --o-visualization weighted-unifrac-pH-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column PFAS \
  --p-pairwise \
  --o-visualization weighted-unifrac-PFAS-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --m-metadata-column Oxygen \
  --p-pairwise \
  --o-visualization weighted-unifrac-oxygen-group-significance.qzv

qiime diversity beta-group-significance \
  --i-distance-matrix core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txtt \
  --m-metadata-column Sugar \
  --p-pairwise \
  --o-visualization weighted-unifrac-sugar-group-significance.qzv
  
```

# Analyze sample composition in the context of continuous metadata

## Create distance matrices from numeric metadata columns

Import metadata file without the starting material samples: "Peskett_Microbiome_Metadata_DistanceMatrix.txt"

Create a distance matrix for each continuous variable column you wish to evaluate: 

```{bash}

qiime metadata distance-matrix \
  --m-metadata-file Peskett_Microbiome_Metadata_DistanceMatrix.txt \
  --m-metadata-column TimePoint \
  --o-distance-matrix distance-matrix-time.qza

## Apply a two-sided Mantel test to identify correlation between two distance matrices
```


### Unweighted UniFrac distances. 

Use --p-intersect_ids to discard mismatches in IDs between matrices and apply the Mantel test to only IDs that are found in both distance matrices.

```{bash}
qiime diversity mantel \
  --i-dm1 core-metrics-results/unweighted_unifrac_distance_matrix.qza \
  --i-dm2 distance-matrix-time.qza \
  --p-intersect-ids \
  --o-visualization mantel-unweighted-Unifrac-time.qzv
```

### Weighted UniFrac

```{bash}
qiime diversity mantel \
  --i-dm1 core-metrics-results/weighted_unifrac_distance_matrix.qza \
  --i-dm2 distance-matrix-time.qza \
  --p-intersect-ids \
  --o-visualization mantel-weighted-Unifrac-time.qzv
```

# Taxonomic Analysis

## Import taxonomy from Ampliseq run

"import" the .tsv file and Convert to .qza:

```{bash}
qiime tools import \
  --type 'FeatureData[Taxonomy]' \
  --input-format TSVTaxonomyFormat \
  --input-path taxonomy.tsv \
  --output-path ampliseq-taxonomy.qza
```

## View the taxonomic composition of the samples with interactive bar plots:

```{bash}
qiime taxa barplot \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --m-metadata-file Peskett_Microbiome_Metadata_NoBlanks.txt \
  --o-visualization taxa-bar-plots.qzv
```

## Create feature tables with taxonomy instead of feature IDs

```{bash}
# Phylum Level
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 2 \
  --o-collapsed-table phyla-table.qza
qiime metadata tabulate \
  --m-input-file phyla-table.qza \
  --o-visualization phyla-table.qzv

# Class 
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 3 \
  --o-collapsed-table class-table.qza 
qiime metadata tabulate \
  --m-input-file class-table.qza \
  --o-visualization class-table.qzv

# Order
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 4 \
  --o-collapsed-table order-table.qza 
qiime metadata tabulate \
  --m-input-file order-table.qza \
  --o-visualization order-table.qzv

# Family
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 5 \
  --o-collapsed-table family-table.qza 
qiime metadata tabulate \
  --m-input-file family-table.qza \
  --o-visualization family-table.qzv

# Genus
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 6 \
  --o-collapsed-table genus-table.qza 
qiime metadata tabulate \
  --m-input-file genus-table.qza \
  --o-visualization genus-table.qzv

# Species
qiime taxa collapse \
  --i-table id-filtered-table.qza \
  --i-taxonomy ampliseq-taxonomy.qza \
  --p-level 7 \
  --o-collapsed-table species-table.qza 
qiime metadata tabulate \
  --m-input-file species-table.qza \
  --o-visualization species-table.qzv
```

Export tables as .tsv files:

```{bash}

#Phylum
qiime tools export \
--input-path phyla-table.qza \
--output-path export/phyla-table

biom convert -i export/phyla-table/feature-table.biom \
-o export/phyla-table/phyla-feature-table.tsv --to-tsv

#Class
qiime tools export \
--input-path class-table.qza \
--output-path export/class-table

biom convert -i export/class-table/feature-table.biom \
-o export/class-table/class-feature-table.tsv --to-tsv

#Order
qiime tools export \
--input-path order-table.qza \
--output-path export/order-table

biom convert -i export/order-table/feature-table.biom \
-o export/order-table/order-feature-table.tsv --to-tsv

#Family
qiime tools export \
--input-path family-table.qza \
--output-path export/family-table

biom convert -i export/family-table/feature-table.biom \
-o export/family-table/family-feature-table.tsv --to-tsv

#Genus
qiime tools export \
--input-path genus-table.qza \
--output-path export/genus-table

biom convert -i export/genus-table/feature-table.biom \
-o export/genus-table/genus-feature-table.tsv --to-tsv

#Species
qiime tools export \
--input-path species-table.qza \
--output-path export/species-table

biom convert -i export/species-table/feature-table.biom \
-o export/species-table/species-feature-table.tsv --to-tsv
```

# Compute relative abundances

```{bash}

# Phylum
qiime feature-table relative-frequency \
--i-table phyla-table.qza \
--o-relative-frequency-table rel-phyla-table.qza
qiime metadata tabulate \
  --m-input-file rel-phyla-table.qza \
  --o-visualization rel-phyla-table.qzv
  
# Class
qiime feature-table relative-frequency \
--i-table class-table.qza \
--o-relative-frequency-table rel-class-table.qza
qiime metadata tabulate \
  --m-input-file rel-class-table.qza \
  --o-visualization rel-class-table.qzv
  
# Order
qiime feature-table relative-frequency \
--i-table order-table.qza \
--o-relative-frequency-table rel-order-table.qza
qiime metadata tabulate \
  --m-input-file rel-order-table.qza \
  --o-visualization rel-order-table.qzv
  
# Family
qiime feature-table relative-frequency \
--i-table family-table.qza \
--o-relative-frequency-table rel-family-table.qza
qiime metadata tabulate \
  --m-input-file rel-family-table.qza \
  --o-visualization rel-family-table.qzv
  
# Genus
qiime feature-table relative-frequency \
--i-table genus-table.qza \
--o-relative-frequency-table rel-genus-table.qza
qiime metadata tabulate \
  --m-input-file rel-genus-table.qza \
  --o-visualization rel-genus-table.qzv

# Species
qiime feature-table relative-frequency \
--i-table species-table.qza \
--o-relative-frequency-table rel-species-table.qza
qiime metadata tabulate \
  --m-input-file rel-species-table.qza \
  --o-visualization rel-species-table.qzv
```

Export tables as .tsv files:

```{bash}

#Phylum
qiime tools export \
--input-path rel-phyla-table.qza \
--output-path export/rel-phyla-table

biom convert -i export/rel-phyla-table/feature-table.biom \
-o export/rel-phyla-table/phyla-rel-table.tsv --to-tsv

#Class
qiime tools export \
--input-path rel-class-table.qza \
--output-path export/rel-class-table

biom convert -i export/rel-class-table/feature-table.biom \
-o export/rel-class-table/class-rel-table.tsv --to-tsv

#Order
qiime tools export \
--input-path rel-order-table.qza \
--output-path export/rel-order-table

biom convert -i export/rel-order-table/feature-table.biom \
-o export/rel-order-table/order-rel-table.tsv --to-tsv

#Family
qiime tools export \
--input-path rel-family-table.qza \
--output-path export/rel-family-table

biom convert -i export/rel-family-table/feature-table.biom \
-o export/rel-family-table/family-rel-table.tsv --to-tsv

#Genus
qiime tools export \
--input-path rel-genus-table.qza \
--output-path export/rel-genus-table

biom convert -i export/rel-genus-table/feature-table.biom \
-o export/rel-genus-table/genus-rel-table.tsv --to-tsv

#Species
qiime tools export \
--input-path rel-species-table.qza \
--output-path export/rel-species-table

biom convert -i export/rel-species-table/feature-table.biom \
-o export/rel-species-table/species-rel-table.tsv --to-tsv
```

# Data Visualizations

Load in the metadata file used in qiime2: 

```{r}
metadata<- read.delim("Peskett_Microbiome_Metadata_NoBlanks.txt", dec = ".")

#Change ID column to sampleID and filter out blanks and NTC

metadata <- metadata %>%
  dplyr::rename(sampleID = ID)
```

## Diversity Figures

Import shannon index qiime artifact

```{r}
shannon <- read_qza("quiime2-Outputs/Diversity/core-metrics-results/shannon_vector.qza")$data %>% rownames_to_column("sampleID") 
```

Check that sampleID matches metadata:

```{r}
gplots::venn(list(metadata=metadata$sampleID, shannon=shannon$sampleID)) 
```

### Unweighted Unifrac PCoA

Load in qiime artifact:

```{r}
unweighted <- read_qza("quiime2-Outputs/Diversity/core-metrics-results/unweighted_unifrac_pcoa_results.qza")
```

Prior to creating static PCoA plots, the proportion of variance explained by axes 1 and 2 in the unweighted unifrac analyses needs to be manually extracted:

```{r}
unweighted_PC1<-signif(unweighted$data$ProportionExplained$PC1, digits = 4)
unweighted_PC2<-signif(unweighted$data$ProportionExplained$PC2, digits = 4)
```

Create an ordination plot using Unweighted UniFrac measurement to visualize beta diversity:

Plot 1 will have participant as the colour and pH as the shape

```{r}
#Create visual theme here and edit as needed for presentation purposes

PCoA_theme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
               panel.background = element_blank(), axis.line = element_line(colour = "black"),
               axis.text.x = element_text(color = "black", size = 14),
               axis.text.y = element_text(color = "black", size = 14),
               axis.title.x = element_text(color = "black", size = 16),
               axis.title.y = element_text(color = "black", size = 16),
               legend.title=element_text(size=16), 
               legend.text=element_text(size=16),
               legend.key = element_rect(fill = "white", color = NA),
               plot.title = element_text(size = 20, face = "bold"),
               panel.border = element_rect(colour = "black", fill=NA, size=1))

#unweighted unifrac beta diversity ordination visualization script
unweighted_ordination <-
  unweighted$data$Vectors %>%
  select(SampleID, PC1, PC2) %>% # different formatting here for sampleID name, convert in next step
  dplyr::rename(sampleID = SampleID) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  dplyr::rename(`Shannon Index` = shannon_entropy) %>% 
  dplyr::rename(`Time Point` = TimePoint)
```

Plot will have Participant as the colour, pH as the shape, TimePoint as the size:

```{r}
# Order the Time Point variable properly:

unweighted_ordination$`Time Point` <- factor(unweighted_ordination$`Time Point`,
                                             levels = c( "Starting_Material", "8", "24", "72"))
# Make Plot:

unweightedplot <- unweighted_ordination %>% 
  ggplot(aes(x=PC1, y=PC2, colour = Participant, shape = pH, size =`Time Point`)) +
  geom_point(alpha= 0.8) + #alpha controls transparency and helps when points are overlapping
  scale_color_brewer(palette = "Dark2") +
    scale_size_discrete(
  labels = c("Starting_Material" = "Starting Material",
             "8" = "8",
             "24" = "24",
             "72" = "72")) +
  scale_shape_discrete(
  labels = c("Starting_Material" = "Starting Material",
             "pH2" = "pH 2",
             "pH7" = "pH 7")) +
  labs(title = "Unweighted UniFrac", 
       x = paste(paste("PC1", unweighted_PC1*100), "%"), y = paste(paste("PC2", unweighted_PC2*100), "%")) +
  PCoA_theme +
  guides(
    colour = guide_legend(order = 1, override.aes = list(size = 4)),
    size   = guide_legend(order = 2),
    shape  = guide_legend(order = 3, override.aes = list(size = 4))) +
  geom_vline(aes(xintercept=0), color="black",
             linetype="dashed") +
  geom_hline(aes(yintercept = 0), color="black",
             linetype="dashed")

unweightedplot
```

#### PFAS versus Controls

```{r}
unweighted_ordination$PFAS <- factor(unweighted_ordination$PFAS,
                                             levels = c( "Starting_Material", "Yes", "No"))

unweightedplot2 <- unweighted_ordination %>% 
  ggplot(aes(x=PC1, y=PC2, colour = PFAS, shape = Participant, size =`Time Point`)) +
  geom_point(alpha= 0.8) + #alpha controls transparency and helps when points are overlapping
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Unweighted UniFrac", 
       x = paste(paste("PC1", unweighted_PC1*100), "%"), y = paste(paste("PC2", unweighted_PC2*100), "%")) +
  PCoA_theme +
  guides(colour = guide_legend(override.aes = list(size=4)),
         shape = guide_legend(override.aes = list(size = 4)))

unweightedplot2
```

### Weighted Unifrac PCoA

Load in qiime artifact:

```{r}
weighted <- read_qza("quiime2-Outputs/Diversity/core-metrics-results/weighted_unifrac_pcoa_results.qza")
```

Prior to creating static PCoA plots, the proportion of variance explained by axes 1 and 2 in the unweighted unifrac analyses needs to be manually extracte:

```{r}
weighted_PC1<-signif(weighted$data$ProportionExplained$PC1, digits = 4)
weighted_PC2<-signif(weighted$data$ProportionExplained$PC2, digits = 4)
```

Create an ordination plot using weighted UniFrac measurement to visualize beta diversity:

```{r}
#weighted unifrac beta diversity ordination visualization script
weighted_ordination <-
  weighted$data$Vectors %>%
  select(SampleID, PC1, PC2) %>% # different formatting here for sampleID name, convert in next step
  dplyr::rename(sampleID = SampleID) %>%
  left_join(metadata) %>%
  left_join(shannon) %>%
  dplyr::rename(`Shannon Index` = shannon_entropy) %>% 
  dplyr::rename(`Time Point` = TimePoint)

weighted_ordination$`Time Point` <- as.factor(unweighted_ordination$`Time Point`)

weighted_ordination$`Time Point` <- factor(weighted_ordination$`Time Point`,
                                             levels = c( "Starting_Material", "8", "24", "72"))

weightedplot <- weighted_ordination %>% 
  ggplot(aes(x=PC1, y=PC2, colour = Participant, shape = pH, size =`Time Point`)) +
  geom_point(alpha= 0.8) + #alpha controls transparency and helps when points are overlapping
  scale_color_brewer(palette = "Dark2") +
  scale_size_discrete(
  labels = c("Starting_Material" = "Starting Material",
             "8" = "8",
             "24" = "24",
             "72" = "72")) +
  scale_shape_discrete(
  labels = c("Starting_Material" = "Starting Material",
             "pH2" = "pH 2",
             "pH7" = "pH 7")) +
  labs(title = "Weighted UniFrac", 
       x = paste(paste("PC1", weighted_PC1*100), "%"), y = paste(paste("PC2", weighted_PC2*100), "%")) +
  PCoA_theme +
  guides(
    colour = guide_legend(order = 1, override.aes = list(size = 4)),
    size   = guide_legend(order = 2),
    shape  = guide_legend(order = 3, override.aes = list(size = 4)))

weightedplot
```

#### PFAS versus Controls

```{r}
weighted_ordination$PFAS <- factor(weighted_ordination$PFAS,
                                             levels = c( "Starting_Material", "Yes", "No"))
weightedplot2 <- weighted_ordination %>% 
  ggplot(aes(x=PC1, y=PC2, colour = PFAS, shape = Participant, size =`Time Point`)) +
  geom_point(alpha= 0.8) + #alpha controls transparency and helps when points are overlapping
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Weighted UniFrac", 
       x = paste(paste("PC1", weighted_PC1*100), "%"), y = paste(paste("PC2", weighted_PC2*100), "%")) +
  PCoA_theme +
  guides(colour = guide_legend(override.aes = list(size=4)),
         shape = guide_legend(override.aes = list(size = 4)))

weightedplot2

```

### Chemistry vectors

#### Unweighted

Prepare data for envfit()
```{r}
#Add chemistry data to abundance data
chemdata2 <- data %>%
  dplyr::rename(sampleID = Sample_ID)

combined_df <- unweighted_ordination %>%
  left_join(chemdata2 %>%
      select(sampleID, PFOA, diPAP, FTOH, PFHxA, monoPAP),
    by = "sampleID")

#Define PFAS columns 

pfas_cols <- c("diPAP", "PFOA", "PFHxA", "monoPAP", "FTOH")

#Select relevant columns and drop NAs

combined_df <- combined_df %>% 
  dplyr::select(sampleID, PC1, PC2, all_of(pfas_cols)) %>% 
  drop_na()

#Use sampleID as rownames for both matrices

rownames(combined_df) <- combined_df$sampleID

ordination_matrix <- combined_df %>% 
  select(PC1, PC2) %>% 
  as.matrix()

pfas_matrix <- combined_df %>% 
  select(all_of(pfas_cols)) %>% 
  as.matrix()

#Check dimensions

stopifnot(nrow(ordination_matrix) == nrow(pfas_matrix))

#Run envfit

fit <- envfit(ordination_matrix, pfas_matrix, permutations = 999)
```

Extract significant vectors from fit:

```{r}
#Extract vector scores and p-values
vector_scores <- scores(fit, display = "vectors")
pvals <- fit$vectors$pvals

#Create data frame with vectors
vector_df <- as.data.frame(vector_scores)
vector_df$pfas <- rownames(vector_df)
vector_df$pval <- pvals

#Keep only significant vectors and scale for better visualization
signif_vectors <- vector_df %>% 
  filter(pval <= 0.05) %>% 
  mutate(PC1 = PC1 * 0.5, PC2 = PC2 * 0.5)
```

Overlay vectors on existing ggplot:
```{r}
unweightedplot_overlay <- unweightedplot +
  geom_segment(data = signif_vectors,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black", size = 1, inherit.aes = FALSE) +
  geom_text(data = signif_vectors,
            aes(x = PC1, y = PC2, label = pfas),
            color = "black", size = 5, inherit.aes = FALSE) +
    geom_vline(aes(xintercept=0), color="black",
             linetype="dashed") +
  geom_hline(aes(yintercept = 0), color="black",
             linetype="dashed")
  
unweightedplot_overlay
```

#### Weighted

Prepare data for envfit()
```{r}
combined_df2 <- weighted_ordination %>%
  left_join(chemdata2 %>%
      select(sampleID, PFOA, diPAP, FTOH, PFHxA, monoPAP),
    by = "sampleID")

#Select relevant columns and drop NAs

combined_df2 <- combined_df2 %>% 
  dplyr::select(sampleID, PC1, PC2, all_of(pfas_cols)) %>% 
  drop_na()

#Use sampleID as rownames for both matrices

rownames(combined_df2) <- combined_df2$sampleID

ordination_matrix2 <- combined_df2 %>% 
  select(PC1, PC2) %>% 
  as.matrix()

pfas_matrix2 <- combined_df2 %>% 
  select(all_of(pfas_cols)) %>% 
  as.matrix()

#Check dimensions

stopifnot(nrow(ordination_matrix2) == nrow(pfas_matrix2))

#Run envfit

fit2 <- envfit(ordination_matrix2, pfas_matrix2, permutations = 999)
```

Extract significant vectors from fit:

```{r}
#Extract vector scores and p-values
vector_scores2 <- scores(fit2, display = "vectors")
pvals2 <- fit2$vectors$pvals

#Create data frame with vectors
vector_df2 <- as.data.frame(vector_scores2)
vector_df2$pfas <- rownames(vector_df2)
vector_df2$pval <- pvals2

#Keep only significant vectors and scale for better visualization
signif_vectors2 <- vector_df2 %>% 
  filter(pval <= 0.05) %>% 
  mutate(PC1 = PC1 * 0.5, PC2 = PC2 * 0.5)
```

Overlay vectors on existing ggplot:
```{r}
weightedplot_overlay <- weightedplot +
  geom_segment(data = signif_vectors2,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black", size = 1, inherit.aes = FALSE) +
  geom_text(data = signif_vectors2,
            aes(x = PC1, y = PC2, label = pfas),
            color = "black", size = 5, inherit.aes = FALSE) +
    geom_vline(aes(xintercept=0), color="black",
             linetype="dashed") +
  geom_hline(aes(yintercept = 0), color="black",
             linetype="dashed")
  
weightedplot_overlay
```

### Combined Figures

```{r}
combine2 <- ggarrange(unweightedplot_overlay, weightedplot_overlay,
                      labels = c("A)", "B)"),
                      common.legend = TRUE, legend = "right")
combine2

```

## Relative-Abundance Heat-map

### Family Levels

```{r}
metadata<- read.delim("Peskett_Microbiome_Metadata_NoBlanks.txt", dec = ".")

#Change ID column to sampleID and filter out blanks and NTC

metadata <- metadata %>%
  dplyr::rename(sampleID = ID)
```

Import shannon index qiime artifact

```{r}
shannon <- read_qza("quiime2-Outputs/Diversity/core-metrics-results/shannon_vector.qza")$data %>% rownames_to_column("sampleID") 
```

```{r}
rel_family_table <- fread("quiime2-Outputs/export/rel-family-table/family-rel-table.tsv", sep = "\t", header = "auto")
rel_family_table2 <- rel_family_table %>% 
                      dplyr::rename(classification = `#OTU ID`) %>%
                      separate(classification, 
                                                     into = c("Kingdom", 
                                                             "Phylum",
                                                             "Class",
                                                             "Order",
                                                             "Family"), sep = ";") %>%
                          
                                            #Phylum
                                            mutate(Phylum = str_replace(string = Phylum,
                                                              pattern = "p__",
                                                              replacement = "")) %>%
                                            mutate(Phylum = str_replace(string = Phylum,
                                                              pattern = "(.*)",
                                                              replacement = "*\\1*")) %>%
                                            #Class
                                            mutate(Class = str_replace(string = Class,
                                                              pattern = "c__",
                                                              replacement = "")) %>%
                                            mutate(Class = str_replace(string = Class,
                                                              pattern = "(.*)",
                                                              replacement = "*\\1*")) %>%
                
                                            #Order
                                            mutate(Order = str_replace(string = Order,
                                                              pattern = "o__",
                                                              replacement = "")) %>%
                                            mutate(Order = str_replace(string = Order,
                                                              pattern = "(.*)",
                                                              replacement = "*\\1*")) %>%
                
                                            #Family
                                            mutate(Family = str_replace(string = Family,
                                                              pattern = "f__",
                                                              replacement = "")) %>%
                                            mutate(Family = str_replace(string = Family,
                                                              pattern = "(.*)",
                                                              replacement = "*\\1*")) %>%

                                            #Unknowns
                                            mutate(Family = str_replace(string = Family, 
                                                            pattern = "__",
                                                            replacement = "*Unidentified family*")) %>%
  
                                            mutate(Phylum = str_replace(string = Phylum, 
                                                            pattern = "__",
                                                            replacement = "*Unidentified phylum*")) %>%
  
                                            #Generate collated labels
                                            mutate(Phylum_family = glue("{Phylum}, {Family}")) %>%
                                            relocate(Phylum_family, .after=Family)
```

Filter out any Eukaryotic and Unassigned sequences and then run the gather function to convert this from a wide to a long data format better suited to data visualization. Use case_when to reproduce the Material names originally present in the metadata file in addition to any other relevant information to facilitate adding additional layers of information in the visualization.

```{r}
#Data clean up and reorganization
rel_family_heatmap <- rel_family_table2 %>%
                        filter(!Kingdom == "Unassigned" & !Kingdom == "d__Eukaryota") %>% #filter out unassigned and eukaryotic sequences
                        gather(key = `sampleID`, value = `Relative abundance`, "B24F10":"MStool", na.rm=FALSE) %>%
                        mutate(`Relative abundance` = `Relative abundance` *100) #convert decimals to percent

# Add in metadata

rel_family_heatmap_meta <- merge(rel_family_heatmap, metadata, by = "sampleID")

rel_family_heatmap_meta2  <- rel_family_heatmap_meta  %>% 
  mutate(
    sampleID2 = sampleID) %>% 
  mutate(sampleID2 = str_replace_all(sampleID, c("F" = "A", "M" = "B")))
```

```{r}
#Create aesthetic theme
heatmap_theme<-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                 panel.background = element_blank(), axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(color = "black", size = 10),
                 axis.text.y = element_text(color = "black", size = 10),
                 axis.title.x = element_text(color = "black", size = 12, vjust = -3),
                 axis.title.y = element_text(color = "black", size = 12, vjust = 6),
                 legend.title=element_text(size=12, vjust = 2), 
                 legend.text=element_text(size=12),
                 legend.key = element_rect(fill = "white", color = NA),
                 plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
                 panel.border = element_rect(colour = "black", fill=NA, size=1),
                 plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))
```

#### PFAS treatment samples only

```{r}
participant_labeller <- as_labeller(c(
  "A" = "Participant A",
  "B" = "Participant B"
))

rel_family_heatmap_meta2$TimePoint <- factor(rel_family_heatmap_meta2$TimePoint, levels = c( "Starting_Material", "8", "24", "72"))

rel_family_heatmap_pfas <- rel_family_heatmap_meta2 %>% 
   mutate(
    pH = as.character(pH),
    TimePoint = as.character(TimePoint)
  )

rel_family_heatmap_pfas <- rel_family_heatmap_pfas %>% 
  mutate(
    pH = ifelse(pH == "Starting_Material", "N/A", pH),
    TimePoint = ifelse(TimePoint == "Starting_Material", "N/A", TimePoint)
  ) %>% 
  filter(PFAS != "No")

rel_family_heatmap_pfas$TimePoint <- factor(rel_family_heatmap_pfas$TimePoint, 
                                            levels = c( "N/A", "8", "24", "72"))

rel_family_heatmap_pfas$sampleID2 <- factor(rel_family_heatmap_pfas$sampleID2,
                                             levels = c( "AStool","B8A1",
                                                         "B8A2",
                                                         "B8A3", "B8A4",
                                                         "B8A5", "B8A6",
                                                         "B8A7", "B8A8",
                                                         "B24A1", "B24A2",
                                                         "B24A3", "B24A4",
                                                         "B24A5", "B24A6",
                                                         "B24A7", "B24A8",
                                                         "B72A1", "B72A2",
                                                         "B72A3", "B72A4",
                                                         "B72A5", "B72A6",
                                                         "B72A7", "B72A8",
                                                         "BStool","B8B1",
                                                         "B8B2",
                                                         "B8B3", "B8B4",
                                                         "B8B5", "B8B6",
                                                         "B8B7", "B8B8",
                                                         "B24B1", "B24B2",
                                                         "B24B3", "B24B4",
                                                         "B24B5", "B24B6",
                                                         "B24B7", "B24B8",
                                                         "B72B1", "B72B2",
                                                         "B72B3", "B72B4",
                                                         "B72B5", "B72B6",
                                                         "B72B7", "B72B8"))

rel_family_heatmap_pfas_only <- rel_family_heatmap_pfas %>%
                                  filter(`Relative abundance` >= 1) %>%     #add cut off value bigger or equal to here 
  ggplot(aes(x= sampleID2, y = reorder(Phylum_family, `Relative abundance`), fill= `Relative abundance`)) +
                                  geom_tile(aes()) + 
                                  facet_nested(~ Participant + pH + TimePoint, 
                                               scales = "free", space = "free", labeller = labeller(Participant = participant_labeller)) +
                                  theme(axis.text.x=element_text(angle=0, hjust=1),
                                  axis.title.y = element_text(color = "black", size = 15, vjust = 0),
                                  legend.key.size = unit(2, 'cm'))+
                                  scale_fill_viridis_c(name="Relative abundance (%)", 
                                                      na.value = "white", limits = c(0, 100)) + #set to max relative abundance rounded up
                                  labs(
                                  x = "Sample ID" , y = "Phylum, family name") +
                                  heatmap_theme +
                                  theme(axis.text.x = element_text(color = "black", size = 10, face = "bold", hjust = 0.5, vjust = 0.2, angle = 90),
                                        axis.text.y = element_markdown(size = 10, face = "bold"),
                                        axis.title.x = element_markdown(color = "black", size = 12, face = "bold"),
                                        axis.title.y = element_text(color = "black", size = 12, face = "bold"),
                                        strip.text.x = element_text(color = "black", size = 10, face = "bold"))

rel_family_heatmap_pfas_only
```

#### Controls version

```{r}
rel_family_heatmap_no_pfas <- rel_family_heatmap_meta2 %>% 
   mutate(
    pH = as.character(pH),
    TimePoint = as.character(TimePoint)
  )

rel_family_heatmap_no_pfas <- rel_family_heatmap_no_pfas %>% 
  mutate(
    pH = ifelse(pH == "Starting_Material", "N/A", pH),
    TimePoint = ifelse(TimePoint == "Starting_Material", "N/A", TimePoint)
  ) %>% 
  filter(PFAS != "Yes")

rel_family_heatmap_no_pfas$TimePoint <- factor(rel_family_heatmap_no_pfas$TimePoint, 
                                            levels = c( "N/A", "8", "24", "72"))

rel_family_heatmap_no_pfas$sampleID2 <- factor(rel_family_heatmap_no_pfas$sampleID2,
                                             levels = c( "AStool","B8A9",
                                                         "B8A10",
                                                         "B8A11", "B8A12",
                                                         "B8A13", "B8A14",
                                                         "B8A15", "B8A16",
                                                         "B24A9", "B24A10",
                                                         "B24A11", "B24A12",
                                                         "B24A13", "B24A14",
                                                         "B24A15", "B24A16",
                                                         "B72A9", "B72A10",
                                                         "B72A11", "B72A12",
                                                         "B72A13", "B72A14",
                                                         "B72A15", "B72A16",
                                                         "BStool","B8B9",
                                                         "B8B10",
                                                         "B8B11", "B8B12",
                                                         "B8B13", "B8B14",
                                                         "B8B15", "B8B16",
                                                         "B24B9", "B24B10",
                                                         "B24B11", "B24B12",
                                                         "B24B13", "B24B14",
                                                         "B24B15", "B24B16",
                                                         "B72B9", "B72B10",
                                                         "B72B11", "B72B12",
                                                         "B72B13", "B72B14",
                                                         "B72B15", "B72B16"))

rel_family_heatmap_no_pfas_plot <- rel_family_heatmap_no_pfas %>%
                                  filter(`Relative abundance` >= 1) %>%     #add cut off value bigger or equal to here 
  ggplot(aes(x= sampleID2, y = reorder(Phylum_family, `Relative abundance`), fill= `Relative abundance`)) +
                                  geom_tile(aes()) + 
                                  facet_nested(~ Participant + pH + TimePoint, 
                                               scales = "free", space = "free", labeller = labeller(Participant = participant_labeller)) +
                                  theme(axis.text.x=element_text(angle=0, hjust=1),
                                  axis.title.y = element_text(color = "black", size = 15, vjust = 0),
                                  legend.key.size = unit(2, 'cm'))+
                                  scale_fill_viridis_c(name="Relative abundance (%)", 
                                                      na.value = "white", limits = c(0, 100)) + #set to max relative abundance rounded up
                                  labs(
                                  x = "Sample ID" , y = "Phylum, family name") +
                                  heatmap_theme +
                                  theme(axis.text.x = element_text(color = "black", size = 10, face = "bold", hjust = 0.5, vjust = 0.2, angle = 90),
                                        axis.text.y = element_markdown(size = 10, face = "bold"),
                                        axis.title.x = element_markdown(color = "black", size = 12, face = "bold"),
                                        axis.title.y = element_text(color = "black", size = 12, face = "bold"),
                                        strip.text.x = element_text(color = "black", size = 10, face = "bold"))

rel_family_heatmap_no_pfas_plot
```

#### Combine

```{r}
combine <- ggarrange(rel_family_heatmap_no_pfas_plot, rel_family_heatmap_pfas_only,
                      labels = c("A) Control", "B) PFAS"), nrow = 2, 
                      heights = c(1,1), common.legend = TRUE, legend = "right")
combine
```

## Chemistry x Microbial Data

### Hierarchical Clustering Correlation Matrix

Construct correlation matrix:

```{r}
#Filter data to only include PFAS samples
family_corr_matrix <- rel_family_heatmap_meta2 %>%
  filter(PFAS == "Yes") %>%
  select(sampleID, sampleID2, Phylum_family, `Relative abundance`,)

#Add chemistry data to abundance data
family_corr_matrix2 <- family_corr_matrix %>%
  left_join(chemdata2 %>%
      select(sampleID, PFOA, diPAP, FTOH, PFHxA, monoPAP),
    by = "sampleID")

# Filter for families that had at least 1 relative abundance measure hit or exceed 1% 
filtered_data <- family_corr_matrix2 %>%
  group_by(Phylum_family) %>%
  filter(any(`Relative abundance` >= 1)) %>%
  ungroup() 

family_corr_matrix2 <- filtered_data %>% 
  pivot_wider(
    names_from = Phylum_family,
    values_from = `Relative abundance`,
    values_fn = mean 
  ) 

#Remove columns with all zeros
family_corr_matrix2 <- family_corr_matrix2[, colSums(family_corr_matrix2 != 0, na.rm = TRUE) > 0]

#Define compound columns 
compound_vars <- c("monoPAP", "PFOA", "diPAP", "FTOH", "PFHxA")

# Extract family names from column names (those that are NOT compound_vars or sampleID)
families <- setdiff( names(family_corr_matrix2),c(compound_vars2, "sampleID", "sampleID2"))

# Initialize correlation matrix
cor_matrix <- matrix(NA, 
                      nrow = length(families), 
                      ncol = length(compound_vars),
                      dimnames = list(families, compound_vars))

# Initialize p-values matrix
p_matrix <- matrix(
  1,
  nrow = length(families),
  ncol = length(compound_vars),
  dimnames = list(families, compound_vars)
)

# Compute correlations
for (fam in families) {
  for (compound in compound_vars) {

    x <- family_corr_matrix2[[fam]]
    y <- family_corr_matrix2[[compound]]

    # Spearman correlation
    cor_matrix[fam, compound] <- cor(
      x, y,
      method = "spearman",
      use = "complete.obs"
    )

    # P-value (safe handling of constant vectors)
    if (length(unique(x)) < 2 || length(unique(y)) < 2) {
      p_matrix[fam, compound] <- 1
    } else {
      p_matrix[fam, compound] <- suppressWarnings(
        cor.test(x, y, method = "spearman", exact = FALSE)$p.value
      )
    }
  }
}

cor_long <- as.data.frame(cor_matrix)
cor_long$Phylum_Family <- rownames(cor_long)
cor_long <- cor_long %>%
  pivot_longer(cols = -Phylum_Family, names_to = "Compound", values_to = "Correlation")

p_long <- as.data.frame(p_matrix)
p_long$Phylum_Family <- rownames(p_long)

p_long <- p_long %>%
  pivot_longer(
    cols = -Phylum_Family,
    names_to = "Compound",
    values_to = "p_value"
  )

#Join p-values to correlation data

cor_long <- cor_long %>%
  left_join(
    p_long,
    by = c("Phylum_Family", "Compound")
  )

#Add FDR correction per family

cor_long <- cor_long %>%
  group_by(Compound) %>%
  mutate(p_fdr = p.adjust(p_value, method = "fdr")) %>%
  ungroup()

#Add stars

cor_long <- cor_long %>%
  mutate(sig = case_when(
    p_fdr < 0.001 ~ "***",
    p_fdr < 0.01  ~ "**",
    p_fdr < 0.05  ~ "*",
    TRUE          ~ ""
  ))

```

Hierarchical clustering:

```{r}
# Convert to matrix with families as rows and compounds as columns:

cor_mat <- cor_long %>%
  select(Phylum_Family, Correlation, Compound) %>% 
  pivot_wider(names_from = Compound, values_from = Correlation) %>%
  column_to_rownames("Phylum_Family") %>%
  as.matrix()

# Create stars matrix

sig_mat <- cor_long %>%
  select(Phylum_Family, sig, Compound) %>% 
  pivot_wider(names_from = Compound, values_from = sig) %>%
  column_to_rownames("Phylum_Family") %>%
  as.matrix()

# Remove asterisks and commas
rownames(cor_mat) <- gsub("[\\*,]", "", rownames(cor_mat))
colnames(cor_mat) <- gsub("[\\*,]", "", colnames(cor_mat))  # optional, if needed
rownames(sig_mat) <- gsub("[\\*,]", "", rownames(sig_mat))
colnames(sig_mat) <- gsub("[\\*,]", "", colnames(sig_mat))

# Generate clustered heat map:

corrplot <- pheatmap(cor_mat,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         color = colorRampPalette(c("#440154", "#21908C", "#FDE725"))(50),
         breaks = seq(-1, 1, length.out = 51),
         border_color = NA,
         display_numbers = sig_mat,
         number_color = "black",
         fontsize_number = 10)

corrplot

```

# Longread Bioinformatics

## Nanophase

```{bash}
conda activate nanophase
```

Run workflow for each sequence:

```{bash}
nanophase meta -l LongRead_B-72-F-16.fastq.gz -t 8 -o Nanophase_B-72-F-16 -e global

nanophase meta -l LongRead_B-72-F-8.fastq.gz -t 40 -o Nanophase_B-72-F-8 -e global

nanophase meta -l LongRead_B-72-M-16.fastq.gz -t 40 -o Nanophase_B-72-M-16 -e global

nanophase meta -l LongRead_B-72-M-8.fastq.gz -t 40 -o Nanophase_B-72-M-8 -e global
```

## De-replicate genomes using dRep

```{bash}
conda activate dRep
```

```{bash}
dRep dereplicate output_directory -g path/to/genomes/*.fasta -comp 70 -con 10 -sa 0.98 -p 20
```

##DRAM annotations

```{bash}
conda activate DRAM_conda
```

Annotate with custom databases included:

```{bash}
DRAM.py annotate -i '/path/to/dereplicated_genomes/*.fasta' \
-o /path/to/outputdirectory \
--use_uniref \
--custom_fasta_loc /path/to/PFASDB_FASTA.fasta \
--custom_db_name PFASDB \
--custom_fasta_loc /path/to/RDaseDB_FASTA.fasta \
--custom_db_name RDaseDB \
--threads 40
```

Run the distill command once the annotation is complete:

```{bash}
DRAM.py distill -i /path/to/annotations.tsv -o /path/to/outputdirectory --trna_path /path/to/trnas.tsv --rrna_path /path/to/rrnas.tsv
```

## Run GTDB-Tk for taxonomic classification and phylogenetic trees

```{bash}
conda activate gtdbtk-2.4.0
```

```{bash}
gtdbtk classify_wf --genome_dir /path/to/dereplicated_genomes --extension fasta --out_dir /path/to/outputdirectory --cpus 40 --pplacer_cpus 40 --skip_ani_screen
```

## Run Sylph

```{bash}
conda activate sylph
```

Sketch single end long reads:

```{bash}
sylph sketch -r *.fastq.gz -d /path/to/Sylph_Sketches -t 40
```

Create a sylph database from the high quality MAGs:

```{bash}
sylph sketch /path/to/dereplicated_genomes/*.fasta -o /path/to/outputdirectory/MAG_sylph_database -t 40 -c 200
```

Download the gtdb-tk database form the Sylph authors:

```{bash}
wget http://faust.compbio.cs.cmu.edu/sylph-stuff/gtdb-r220-c200-dbv1.syldb
```

Run Sylph:

```{bash}
sylph profile /path/to/MAG_sylph_database.syldb /path/to/Sylph_sketches_directory/*.sylsp  -t 40 -o /path/to/output_directory/sylph_results.tsv
```

## SqueezeMeta

```{bash}
conda activate SqueezeMeta_dev
```

When running in sequential mode you can specify the external assembly file for each sample in the sample file.

Example:

Sample1 readfileD_1.fastq pair1 extassembly=/path/to/extassembly 

Need to provide the proper mapper - minimap2 
Using minimap2-ont for Oxford Nanopore Technologies (ONT) reads

```{bash}
SqueezeMeta.pl \
-m sequential \
-s /path/to/SampleList.txt \
-f /path/to/raw_fastq_directory \
-c 1000 \
-map minimap2-ont \
-binners concoct,metabat2,maxbin \
-extdb /path/to/custom_database_list.txt \
-t 40
```

Then, will want to re-run the functional annotation portion with a minimum identity value of 50:

Project names: B-72-F-8, B-72-F-16, B-72-M-8, B-72-M-16

Started with project name: B-72-F-16

Edited the parameters.pl file to increase the minimum identity from 30% to 50%.

Editing the parameters.pl files:

```{bash}
#Make a back-up copy of each file

cp parameters.pl parameters_backup.pl

#open file

vi parameters.pl 

#Steps for editing:

#Press i to enter Insert mode.
#Use the arrow keys to make changes (e.g., $DB_NAME = 'newdb';).
#Press Esc to exit Insert mode.
#Type :wq and press Enter to save and quit.
```

After you finish editing the two files restart the SqueezeMeta run at Step 7. Functional Assignment:

```{bash}

#B-72-F-16

SqueezeMeta.pl \
-p B-72-F-16 \
--restart \
-step 7 \
--force_overwrite

#B-72-F-8

SqueezeMeta.pl \
-p B-72-F-8 \
--restart \
-step 7 \
--force_overwrite

#B-72-M-16

SqueezeMeta.pl \
-p B-72-M-16 \
--restart \
-step 7 \
--force_overwrite

#B-72-M-8

SqueezeMeta.pl \
-p B-72-M-8 \
--restart \
-step 7 \
--force_overwrite

# Create Zip files using sqm2zip.py
```


# Longread Visualizations

## Relative abundance heat map

### Species/bin level

```{r}
#Create aesthetic theme
heatmap_theme<-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
                 panel.background = element_blank(), axis.line = element_line(colour = "black"),
                 axis.text.x = element_text(color = "black", size = 12),
                 axis.text.y = element_text(color = "black", size = 12),
                 axis.title.x = element_text(color = "black", size = 15, vjust = -3, face = "bold"),
                 axis.title.y = element_text(color = "black", size = 15, vjust = 6, face = "bold"),
                 legend.title=element_text(size=15, vjust = 2), 
                 legend.text=element_text(size=12),
                 legend.key = element_rect(fill = "white", color = NA),
                 plot.title = element_text(size = 15, face = "bold", hjust = 0.5),
                 panel.border = element_rect(colour = "black", fill=NA, size=1),
                 plot.margin = unit(c(2, 2, 2, 2), "cm"))
```


```{r}
MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff <-  Sylph_GTDB_outputs_combined %>%
                                                    ggplot(aes(x= Sample_ID, y = reorder(Species_bin, Taxonomic_abundance), fill= Taxonomic_abundance)) +
                                                    geom_tile(aes()) + 
                                                    facet_nested(~Treatment + Participant, 
                                                                 scales = "free", space = "free", labeller = labeller(Participant = participant_labeller)) +
                                                    theme(axis.text.x=element_text(angle=0, hjust=1),
                                                    axis.title.y = element_text(color = "black", size = 12, vjust = 0),
                                                    legend.key.size = unit(2, 'cm'))+
                                                    scale_fill_viridis_c(name="Relative abundance (%)", 
                                                                        na.value = "white", limits = c(0, 95)) + #set to max relative abundance rounded up
                                                    labs(
                                                    x = "Sample ID" , y = "MAG ID") +
                                                    heatmap_theme +
                                                    theme(axis.text.x = element_text(color = "black", size = 12, face = "bold", hjust = 0.5, angle = 0),
                                                          axis.text.y = element_markdown(size = 12, face = "bold"),
                                                          axis.title.x = element_text(color = "black", size = 15, face = "bold"),
                                                          strip.text.x = element_text(color = "black", size = 15, face = "bold"))


MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff
```

### Family level

```{r}
family_abundance <- Sylph_GTDB_outputs_combined %>%
  group_by(Sample_ID, Treatment, Participant, Family) %>%
  summarise(
    Taxonomic_abundance = sum(Taxonomic_abundance, na.rm = TRUE),
    .groups = "drop"
  )

MAG_taxonomic_abundance_heatmap_family_no_cutoff <-  family_abundance %>%
                                                    ggplot(aes(x= Sample_ID, y = reorder(Family, Taxonomic_abundance), fill= Taxonomic_abundance)) +
                                                    geom_tile(aes()) + 
                                                    facet_nested(~Treatment + Participant, 
                                                                 scales = "free", space = "free", labeller = labeller(Participant = participant_labeller)) +
                                                    theme(axis.text.x=element_text(angle=0, hjust=1),
                                                    axis.title.y = element_text(color = "black", size = 12, vjust = 0),
                                                    legend.key.size = unit(2, 'cm'))+
                                                    scale_fill_viridis_c(name="Relative abundance (%)", 
                                                                        na.value = "white", limits = c(0, 100)) + #set to max relative abundance rounded up
                                                    labs(
                                                    x = "Sample ID" , y = "Family name") +
                                                    heatmap_theme +
                                                    theme(axis.text.x = element_text(color = "black", size = 12, face = "bold", hjust = 0.5, angle = 0),
                                                          axis.text.y = element_markdown(size = 12, face = "bold"),
                                                          axis.title.x = element_text(color = "black", size = 15, face = "bold"),
                                                          strip.text.x = element_text(color = "black", size = 15, face = "bold"))


MAG_taxonomic_abundance_heatmap_family_no_cutoff

```

## PFAS Genes Presence/Absence

Load in DRAM annotations
```{r}
DRAM_annotations <- fread("annotations.tsv", sep = "\t", header = "auto") #68921 observations
head(DRAM_annotations)

DRAM_annotations<- DRAM_annotations %>% 
  dplyr::rename(genome = fasta) 
```

Filter PFASDB using some more strict cut-offs:

```{r}
Quality_PFASDB_annotations <- DRAM_annotations %>%
                                      filter(PFASDB_identity >= 0.5 & PFASDB_eVal <= 1e-3 & PFASDB_bitScore >= 100) #148 observations
```

Remove ID from the PFAS gene label in order to group genes into categories

Load in annotation summary info to add short form for the genes and added a gene column to the DRAM annotation file:

```{r}
PFASDB_summary <- read.csv("PFASDB_merged_Annotation_Info.csv")

Quality_PFASDB_annotations_categories <-
  Quality_PFASDB_annotations %>%
   mutate(Gene_name = str_remove(PFASDB_hit, fixed(PFASDB_id))) %>% 
  mutate(Entry = str_remove(PFASDB_id, "^.*_")) %>% 
  left_join(
    PFASDB_summary %>% select(Entry, Gene),
    by = "Entry"
  ) %>%
  dplyr::rename(PFAS_gene = Gene)
```

Load in DRAM product and connect to Sylph and GTDB combined dataframe:

```{r}
DRAM_product <- fread("product.tsv", sep = "\t", header = "auto") 

DRAM_product_joined <- left_join(Sylph_GTDB_outputs_combined, DRAM_product, by = "genome")
```

Left join the taxonomy and relative abundance data
```{r}

DRAM_product_PFAS_genomes <- DRAM_product_joined %>%
                         left_join(Quality_PFASDB_annotations_categories) %>% 
                                mutate(PFAS_gene_present = if_else(!is.na(PFAS_gene),
                                                                   "Present", "Absent")) %>%
                                relocate(c(Taxonomic_abundance, PFAS_gene, PFAS_gene_present),
                                         .after = "genome") #relocate the taxonomy, relative abundance, and PFAS gene data to the first columns for ease of reference
```

Plot:

```{r}
PFAS_gene_presence_absence_heatmap <-  DRAM_product_PFAS_genomes %>%
  mutate(Species_bin_ordered = reorder(Species_bin, Taxonomic_abundance)) %>%
  tidyr::complete(
    Species_bin_ordered,
    PFAS_gene,
    fill = list(PFAS_gene_present = NA)
  ) %>%
  ggplot(aes(x = PFAS_gene, y = Species_bin_ordered, fill = PFAS_gene_present)) +
  geom_tile(aes()) + 
  scale_fill_manual(
    name = "Gene presence",
    values = c("Present" = "darkgreen", "Absent" = "turquoise"),
    na.value = "turquoise", breaks = c("Present", "Absent")) + 
  scale_x_discrete(na.translate = FALSE) + #Keep NAs in data, but don't show them on the x-axis
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  labs(x = "", y = "MAG ID") +
  heatmap_theme +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_markdown(size = 15, face = "bold")
  )

PFAS_gene_presence_absence_heatmap
```

Join the relative abundance and the PFAS gene presence/absence data, preserving both y-axes initially to double check organization prior to making final figure:

```{r}
PFAS_gene_presence_absence_heatmap_no_y <-  PFAS_gene_presence_absence_heatmap +
                                                                theme(axis.text.y = element_blank(),
                                                                      axis.title.y = element_blank())

MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff_no_title <- MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff + 
                                                            theme(plot.title = element_blank()) 

# Left plot: MAG taxonomic abundance
MAG_left <- MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff_no_title +
  theme(
    plot.margin = margin(t = 5, r = 2, b = 50, l = 40, unit = "pt")  # increase bottom margin
  )


# Right plot: PFAS gene presence/absence
PFAS_right <- PFAS_gene_presence_absence_heatmap_no_y +
  labs(x = "Gene") +  # ensure x-axis title is set
  theme(
    axis.text.y = element_blank(),       # hide y-axis labels
    axis.title.y = element_blank(),      # hide y-axis title
    plot.margin = margin(t = 5, r = 5, b = 50, l = 2, unit = "pt"),  # enough space for x-axis title
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12),
    axis.title.x = element_text(size = 15, face = "bold")  # style the title
  ) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

# Combine with patchwork
combined_heatmap <- MAG_left + PFAS_right +
  plot_layout(ncol = 2, widths = c(2, 1), guides = 'collect') &
  theme(legend.position = "right")

combined_heatmap
```

## Conjugation Genes

```{r}
# Filter for conjugation - related genes
conjugation <- DRAM_annotations %>% 
  filter(ko_id %in% c("K01023", "K00799", "K12999")) %>%
  mutate(conjugation_gene_present = "Present")

#join with DRAM porduct
DRAM_product_conjugation_genomes <- DRAM_product_joined %>%
  left_join(
    conjugation %>% 
      select(genome, ko_id, conjugation_gene_present),
    by = "genome"
  ) %>% 
  mutate(
    conjugation_gene_present = replace_na(
      conjugation_gene_present, "Absent"
    )
  ) %>% 
  relocate(
    c(Taxonomic_abundance, ko_id, conjugation_gene_present),
    .after = "genome"
  )
```

Plot:

```{r}
conjugation_gene_presence_absence_heatmap <-  DRAM_product_conjugation_genomes %>%
  mutate(Species_bin_ordered = reorder(Species_bin, Taxonomic_abundance)) %>%
  tidyr::complete(
    Species_bin_ordered,
    ko_id,
    fill = list(conjugation_gene_present = NA)
  ) %>%
  ggplot(aes(x = ko_id, y = Species_bin_ordered, fill = conjugation_gene_present)) +
  geom_tile(aes()) + 
  scale_fill_manual(
    name = "Gene presence",
    values = c("Present" = "darkgreen", "Absent" = "turquoise"),
    na.value = "turquoise", breaks = c("Present", "Absent")) + #Keep NAs in data, but don't show them on the x-axis
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  labs(x = "", y = "MAG ID") +
  heatmap_theme +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_markdown(size = 15, face = "bold")
  ) + 
  scale_x_discrete(na.translate = FALSE) 

conjugation_gene_presence_absence_heatmap
```

## ALP gene

```{r}
# Filter for ap - related genes
alp <- DRAM_annotations %>% 
  filter(ko_id %in% c("K01077")) %>%
  mutate(alp_gene_present = "Present")

#join with DRAM porduct
DRAM_product_alp_genomes <- DRAM_product_joined %>%
  left_join(
    alp %>% 
      select(genome, ko_id, alp_gene_present),
    by = "genome"
  ) %>% 
  mutate(
    alp_gene_present = replace_na(
      alp_gene_present, "Absent"
    )
  ) %>% 
  relocate(
    c(Taxonomic_abundance, ko_id, alp_gene_present),
    .after = "genome"
  )
```

Plot:

```{r}
alp_gene_presence_absence_heatmap <-  DRAM_product_alp_genomes %>%
  mutate(Species_bin_ordered = reorder(Species_bin, Taxonomic_abundance)) %>%
  tidyr::complete(
    Species_bin_ordered,
    ko_id,
    fill = list(alp_gene_present = NA)
  ) %>%
  ggplot(aes(x = ko_id, y = Species_bin_ordered, fill = alp_gene_present)) +
  geom_tile(aes()) + 
  scale_fill_manual(
    name = "Gene presence",
    values = c("Present" = "darkgreen", "Absent" = "turquoise"),
    na.value = "turquoise", breaks = c("Present", "Absent")) + #Keep NAs in data, but don't show them on the x-axis
  scale_x_discrete(drop = FALSE) +
  scale_y_discrete(drop = FALSE) +
  labs(x = "", y = "MAG ID") +
  heatmap_theme +
  theme(
    axis.text.x = element_text(size = 15, face = "bold", angle = 90, hjust = 1, vjust = 0.5),
    axis.text.y = element_markdown(size = 15, face = "bold")
  ) + 
  scale_x_discrete(na.translate = FALSE) 

alp_gene_presence_absence_heatmap
```


## Combine conjugation and AP plots

Join the relative abundance and the PFAS gene presence/absence data, preserving both y-axes initially to double check organization prior to making final figure.

```{r}
conjugation_presence_absence_heatmap_no_y <-  conjugation_gene_presence_absence_heatmap +
                                                                theme(axis.text.y = element_blank(),
                                                                      axis.title.y = element_blank())
conjugation_presence_absence_heatmap_no_y

# Left plot: MAG taxonomic abundance
MAG_left <- MAG_taxonomic_abundance_heatmap_species_bin_no_cutoff_no_title +
  theme(
    plot.margin = margin(t = 5, r = 2, b = 50, l = 40, unit = "pt")  # increase bottom margin
  )

# Right plot: gene presence/absence
conjugation_right <- conjugation_presence_absence_heatmap_no_y +
  labs(x = "KEGG ID") +  # ensure x-axis title is set
  theme(
    axis.text.y = element_blank(),       # hide y-axis labels
    axis.title.y = element_blank(),      # hide y-axis title
    plot.margin = margin(t = 5, r = 5, b = 50, l = 2, unit = "pt"),  # enough space for x-axis title
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12),
    axis.title.x = element_text(size = 15, face = "bold")  # style the title
  ) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  scale_x_discrete(na.translate = FALSE) 

# Third plot: alkaline phosphatase gene presence/absence
alp_right <- alp_gene_presence_absence_heatmap +
  labs(x = NULL) +   # remove KEGG ID label
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(t = 5, r = 5, b = 50, l = 2, unit = "pt"),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12)
  ) +
  scale_x_discrete(
    labels = function(x) rep("Alkaline phosphatase", length(x)),
    na.translate = FALSE
  )

# Combine all three
combined_heatmap2 <- MAG_left + conjugation_right + alp_right +
  plot_layout(
    ncol = 3,
    widths = c(2, 0.75, 0.25),   # AP panel narrower than conjugation
    guides = "collect"
  ) &
  theme(legend.position = "right",
        legend.margin = margin(t = 20, r = 5, b = 5, l = 5))

combined_heatmap2
```

# SqueezeMeta

Load in combined SQM object:

```{r}
Squeezemeta_combined_data <- readRDS("Squeezemeta_combined_data.rds")
```

Change the sample IDs to reflect the Participant labels:

```{r}
Squeezemeta_combined_data2 <- Squeezemeta_combined_data

#Abundance
#See what the column names are currently:

colnames(Squeezemeta_combined_data2$taxa$family$abund) 

# Change column names:

colnames(Squeezemeta_combined_data2$taxa$family$abund) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

colnames(Squeezemeta_combined_data2$taxa$family$abund)

#Percent
#See what the column names are currently:

colnames(Squeezemeta_combined_data2$taxa$family$percent) 

# Change column names:

colnames(Squeezemeta_combined_data2$taxa$family$percent) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

```

### Taxonomy

Plot taxanomy and divide based on PFAS and Control

```{r}
taxplot <- plotTaxonomy(Squeezemeta_combined_data2, rank = 'family', count ='percent', metadata_groups = list('Control' = c('B-72-A-16','B-72-B-16'),'PFAS' = c('B-72-A-8','B-72-B-8')))

taxplot
```

### Functional annotations

```{r}
# KEGG

colnames(Squeezemeta_combined_data2$functions$KEGG$cpm) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

KEGGplot <- plotFunctions(Squeezemeta_combined_data2, fun_level = 'KEGG',  count = 'cpm', metadata_groups = list('Control' = c('B-72-A-16','B-72-B-16'),'PFAS' = c('B-72-A-8','B-72-B-8')))
  
KEGGplot

# RDaseDB

colnames(Squeezemeta_combined_data2$functions$RDaseDB$cpm) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

RDaseplot <- plotFunctions(Squeezemeta_combined_data2, fun_level = 'RDaseDB', N = 10, count = 'cpm', metadata_groups = list('Control' = c('B-72-A-16','B-72-B-16'),'PFAS' = c('B-72-A-8','B-72-B-8')))
  
RDaseplot

# PFASDB

#cpm

colnames(Squeezemeta_combined_data2$functions$PFASDB$cpm) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

PFASDBplot <- plotFunctions(Squeezemeta_combined_data2, fun_level = 'PFASDB', count = 'cpm', metadata_groups = list('Control' = c('B-72-A-16','B-72-B-16'),'PFAS' = c('B-72-A-8','B-72-B-8')))
  
PFASDBplot
```

### Subset functions:

```{r}
Squeezemeta_combined_data_benD = subsetFun(Squeezemeta_combined_data2, "_benD_", fixed = FALSE)

colnames(Squeezemeta_combined_data_benD$functions$PFASDB$cpm) <- c("B-72-A-8", "B-72-A-16", "B-72-B-8", "B-72-B-16")

benDplot <- plotFunctions(Squeezemeta_combined_data_benD, fun_level = 'PFASDB', count = 'cpm', metadata_groups = list('Control' = c('B-72-A-16','B-72-B-16'),'PFAS' = c('B-72-A-8','B-72-B-8')))

benDplot
```
